|

### 2024-09-06 manuscript/paperCMSACversion.Rmd
updarted.

@Silverwood2015 looked anecdotally at how enforcers fit into a team atmosphere noticed that teammates put their enforcers on a pedestal. Performing well as an enforcer leads to the player being seen as a protector. When an enforcer fulfills these additional responsibilities to being on the ice as a normal player, players say that it is common for that enforcer to receive perks, such as not paying for their own meals or getting the best seat on the bus. They go as far as to say these players are "deified", emphasizing the social aspects of hockey that surround this history of violence.

While any player on a hockey team can get into a fight, enforcers are exposed to many more fights than the typical NHL player. As a result of being exposed to additional contact, they are potentially exposed to additional head trauma beyond the amount associated with playing hockey. There are many questions about the long-term implications of this head trauma and its potential to result in negative health outcomes in the future. The ingrained nature of fighting into the game of hockey is greatly criticized and there are many active debates about its current declining trend. Figure \ref{fig:enforcer_density} shows a decline in the proportion of enforcers to non-enforcers as players  were born more recently. The question is, should the game continue to move away from contact-driven strategy in favor of younger, faster, highly-skilled players?

```{r enforcer_density, echo = FALSE, fig.cap='Density Plot of Study Cohorts by DOB'}
cohort_dat = dat %>% mutate(cohort = case_when(PIM.GP > 3 ~ 'Enforcer',
                                  #Total.Fights > 50 ~ 'Enforcer',
                                  .default = 'Non-Enforcer'))

@@

Recently, @Popkin2023 explored the effects of being an enforcer on mortality rates, ages at death, and causes of death to address the long-term impact this repeated contact might have. They defined enforcers in two ways: (1) those who had 50 or more career fights (E-F cohort) and (2) those who had, on average, more than 3 penalty minutes per game (E-P cohort).

@@

As mentioned earlier, the outcome of this study is relevant to the debates about regulations and discourse around head trauma in hockey. An accurate analysis is imperative to prevent the spread of misinformed conclusions and best inform regulatory safety practices going forward.

@@

>>>>

@Silverwood2015, who looked anecdotally at how enforcers fit into a team atmosphere,  noticed that teammates put their enforcers on a pedestal. Performing well as an enforcer leads to the player being seen as a protector. When an enforcer fulfills these additional responsibilities to being on the ice as a normal player, players say that it is common for that enforcer to receive perks, such as not paying for their own meals or getting the best seat on the bus. They go as far as to say these players are "deified", emphasizing the social aspects of hockey that surround this history of violence.

While any player on a hockey team can get into a fight, enforcers are exposed to many more fights than the typical NHL player. As a result of being exposed to additional contact, they are potentially exposed to additional head trauma beyond the amount associated with playing hockey. There are many questions about the long-term implications of this head trauma and its potential to result in negative health outcomes in the future. The ingrained nature of fighting into the game of hockey is greatly criticized, and there are many active debates about its current declining trend. Figure \ref{fig:enforcer_density} shows a decline in the proportion of players in the league who are considered enforcers (as defined by @Popkin2023) among players were born more recently. 
<!-- The question is, should the game continue to move away from contact-driven strategy in favor of younger, faster, highly-skilled players? -->

```{r enforcer_density, echo = FALSE, fig.cap='Density Plot of Study Cohorts by DOB'}
#Note: I think the blue should be labeled as total players because it's stacked 
cohort_dat = dat %>% mutate(cohort = case_when(PIM.GP > 3 ~ 'Enforcer',
                                  #Total.Fights > 50 ~ 'Enforcer',
                                  .default = 'Non-Enforcer'))

@@

@Popkin2023 explored the effects of being an enforcer on mortality rates, ages at death, and causes of death to address the long-term impact this repeated contact might have. They defined enforcers in two ways: (1) those who had 50 or more career fights (E-F cohort) and (2) those who had, on average, more than 3 penalty minutes per game (E-P cohort).

@@

The outcome of this study is relevant to the debates about regulations and discourse around head trauma in hockey (and other sports).

An accurate analysis is imperative to prevent the spread of misinformed conclusions and best inform regulatory safety practices going forward.  There is much work demonstrating that elite athletes on average live longer than the general population. @Antero2021, as just one example, estimates that "Olympic athletes live 5.1 years longer than the general population" and @Garatachea2014 presents a large meta-analysis of elite-level athletes life spans and concludes that life expectancy for these elite-level athletes is higher when compared to the population at large.  
However, several studies that have examined life expectancy for athletes involved in sports with more physical contact/ more collisions have found evidence of associations between this contact and reduced life expectancy.  @Zwiers2012 examined Olympians across many sports and found that those elite athletes participating in sports with high levels of physical contact exhibited mortality at higher rates later in life than Olympians in less contact heavy sports.  They then estimate the hazard ratio associated with high contact sports to be 1.13 (95% CI: 1.06-1.21).  @Bianco2007 looked at life expectancy across different sports and did not find any significant results.  However, they note that the median life expectancy for boxers was 73 years compared to an overall median of 76 years.

<!-- https://journals.lww.com/headtraumarehab/fulltext/2012/11000/Life_Expectancy_Following_Rehabilitation__A_NIDRR.15.aspx -->

<!-- Bianco2007:  "Median LE of the samples was 76.0 yrs and no differences were observed in different sports, although it was lower in boxers (73.0 yrs) and higher in tennis players (79.0 yrs)."  -->

<!-- "Those who engaged in disciplines with high levels of physical contact had higher mortality than other Olympians later in life." "the increased mortality associated with high physical contact persisted (hazard ratio 1.13, 1.06 to 1.21)" -->

<!-- However, contact bla h blahj -->

<!-- : The evidence available indicates that top-level athletes live longer than the general population and have a lower risk of 2 major causes of mortality, namely, CVD and cancer. -->

<!-- Antero2021: Olympic athletes live 5.1 years longer than the general population. -->

<!-- Zwiers2012: "Those who engaged in disciplines with high levels of physical contact had higher mortality than other Olympians later in life." "the increased mortality associated with high physical contact persisted (hazard ratio 1.13, 1.06 to 1.21)" -->

@@

Talk about continuous penality minutes stuff?
